---
title: "data Viz real"
author: "Allie Cole"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use packages:
library(tidyverse)
library(here)
library(janitor)
library(lubridate)

# Specifically for plots:
library(patchwork)
library(ggrepel)  #replusive labels
library(gghighlight) 
library(paletteer) #for colors 
library(ggExtra) #for marginal plots
library(ggbeeswarm) #for beeswarm plots

# Some more data to look at
library(gapminder)

# Spatial
library(sf)
```


```{r}
# get the lizards data 

lizards <- read_csv(here("data_tidy", "lizards.csv"))

```

### Make a histogram (really bad one)

```{r}
#lizard weights

ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "blue", 
                 color = "green", 
                 size = 1.2, #borderline
                 linetype = "dotdash")
```


### Make a scatter plot 

```{r}
#length on x and total weight on y 

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 123, 
             fill = "green", 
             color = "purple", 
             alpha = 0.5, 
             size = 4)
```


### Make a bubble plot 


```{r}
# color of point changes based on common name and size vaires based on total length 

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, 
                 size = total_length),
             shape = 22, fill = "light green") +
  theme_minimal()

```

### Facet_grid v Facet_wrap



```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_minimal() +
  facet_wrap(~common_name)
```


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_minimal() +
  facet_grid(sex ~ tail)
```

### Getting things in order!

```{r}
#lizard count by common name 

lizard_count <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# same thing 

##lizard_count <- lizards %>% count(common_name)



ggplot(data = lizard_count, 
       aes(y = fct_reorder(common_name, count), 
                                x = count)) + 
         geom_col(aes(color = common_name, fill = common_name)) +
        labs(x = "Lizard Counts",
             y = "Common Name"
       )
  
```


Try converting commonname to an ordered factor outside of ggplot 

```{r}
common_name_or <- lizard_count %>% 
  mutate(common_name = fct_reorder(common_name, count))
```


### Playing with scales!
#### Axis cale breaks, limits and labels 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 100, 250, 500), 
                     limits = c(0, 500), 
                     expand = c(0, 0)) #gets rid of the space to the side
```


```{r}
lizards_date <- lizards %>% 
  mutate(date = mdy(date)) %>% 
  group_by(date) %>% 
  summarise(count = n())


ggplot(data = lizards_date, aes(x = date, y = count)) +
  geom_line() + 
  scale_x_date(date_breaks ="3 years", 
               date_labels = "%Y")
```


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("purple", "green", "blue", "pink"), 
                     breaks = c(0, 20, 40, 50))
```

### Update a color scheme using a palette in paletter 
make a boxplot with common name on y and total lenth on teh x with color changing based on common name

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, .x = total_length, .fun = median))

ggplot(data = lizards_fct, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "colorBlindness::ModifiedSpectralScheme11Steps")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.background = element_rect(fill = "light yellow", 
                                    color = "purple", 
                                    size = 5), 
    panel.grid.major.x = element_line(color = "green"), 
    panel.grid.minor.x = element_line(color = "light green"), 
    panel.grid.major.y = element_line(color = "blue"), 
    panel.grid.minor.y = element_line(color = "light blue"), 
    plot.background = element_rect(fill = "pink"), 
    axis.text.x = element_text(color = "blue"), 
    axis.text.y = element_text(color = "green"), 
    axis.title = element_text(color = "red")
  )
```

## DAY 9 CONTINUATION of DATA VIZ

### ggrepel 

make a subset from lizards called ww_lizards that only contains western whiptails from site sand 

```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")


ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() + 
  geom_text_repel(aes(label = toe_num), size = 4)

```

```{r}
gapminder <- gapminder %>% 
  filter(continent == "Europe", year == "1972")

# filter for  eurpe create a plot of gdp per capita o x and life expecnecy on teh y labeled by country nam e

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + 
  geom_text_repel(aes(label = country), overlap = 10)

```


### gghighlight 

```{r}
p <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name))


p + gghighlight(toe_num == 250, label_key = toe_num)
```

```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) + 
  gghighlight(max(weight > 30), label_key = common_name)

q
```


### Patchwork for compound figures 
so its easier to put multiple figures together 

```{r}

(((p | q ) / q ) + q * p ) /q & theme_classic()

```

## Explore new graph types 


### Marginal Plots (ggextra)

```{r}
whiptail <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight) 

ggplot(data = whiptail, aes(x = total_length, y = weight)) + 
  geom_point() + 
  geom_rug()
```

marginal blot with box plots 

```{r}
plot <- ggplot(data = whiptail, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = sex), size = 2) + 
  scale_color_manual(values = c("cyan4", "goldenrod", "green"), 
                     name = "Sex", 
                     labels = c("female", "juvenile", "male")) + 
    theme_classic() + 
  theme(legend.position = "bottom")

ggMarginal(plot, type = "boxplot", groupColour = TRUE)
```


### beeswarm plots

```{r}
ggplot(data = whiptail, aes(x = sex, y = weight)) + 
  geom_beeswarm() + 
  geom_violin(fill = NA) + 
  geom_boxplot(fill = NA)
```

### HEat map with geom_tile() 

make a new data frame called lizars counts with date convered to class date and  then count by year and common name the number of lizards observed
```{r}
lizard_count <- lizards %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>% 
  summarize(count = n())


ggplot(data = lizard_count, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count)) + 
  geom_text(aes(label = count), color = "white", size = 2) + 
  scale_fill_viridis_c() #can also do the bottom part, this one is colorblind friendly 
  scale_fill_gradientn(colors = c("goldenrod", "cyan4", "green", "purple", ))
```

### Gonna make a map!

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()


ggplot(data = jornada_veg) + 
  geom_sf(aes(fill = name), color = "white", size = 0.1) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") + 
  labs(x = "Longitude", 
       y = "Latitude", 
       fill = "Dominnt Vegatation") 
  
```










